Composable Generalized Consensus

### 2016-01-06 paper/abstract3.tex
Switched to the llncs template. Reworked the abstract and intro.

\abstract{
State-Machine Replication (SMR) is a basic building block of reliable
distributed systems. Improving SMR algorithms used by cloud-computing providers
can have a big impact on the reliability of the many computing services they
support. However, doing so is hard, often justifying a publication in a top conference.

@@

The key novelty of our framework is its wide applicability: we show that
a large class of SMR algorithms can be expressed in our framework by proposing
two specifications capturing the essence of modular SMR algorithms based on quorum systems. We
demonstrate that concrete algorithms can easily be obtained by refinement from
our abstract specifications and we show that our framework enables formal specification and verification of composite SMR algorithms in TLA+ and Isabelle/HOL.
}

>>>>

\begin{abstract}
State-Machine Replication (SMR) is a basic building block of reliable
distributed systems. Improving SMR algorithms used by cloud-computing providers
can have a big impact on the reliability of the many computing services they
support. However, doing so is hard, often justifying a publication in a top
conference.

@@

The key novelty of our framework is its wide applicability: we show that a
large class of SMR algorithms can be expressed in our framework by proposing
two specifications capturing the essence of modular SMR algorithms based on
quorum systems. We demonstrate that concrete algorithms can easily be obtained
by refinement from our abstract specifications and we show that our framework
enables formal specification and verification of composite SMR algorithms in
TLA+ and Isabelle/HOL\@.
\end{abstract}

### 2016-01-06 paper/intro.tex
Switched to the llncs template. Reworked the abstract and intro.

The State-Machine Replication technique (abbreviated SMR) allows building
reliable services on top of unreliable hardware. The basic idea is to
replicate a service over several servers, each executing the same sequence of
deterministic commands. In such a setup, any replica can be used to determine
the result of a command. Therefore, if a replica crashes, clients can be
transparently redirected to another one and will have the illusion of accessing
a centralized service at all times.

Making SMR algorithms efficient has been a long-standing challenge
and a multitude of algorithms have been proposed. Examples include
Paxos \cite{lamport2001paxos}, Fast Paxos \cite{Lamport06FastPaxos},
Disk Paxos \cite{GafniLamport03DiskPaxos}, Chain Replication
\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability},
Ring Paxos \cite{MarandiETAL10RingPaxosHighthroughputAtomicBroadcastProtocol},
Egalitarian Paxos
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments},
Multi-Coordinated Paxos \cite{CamargosSchmidtPedone07MulticoordinatedPaxos},
Cheap Paxos \cite{LamportMassa04CheapPaxos}, Vertical Paxos
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, Paxos-MIC
\cite{HurfinMoiseNarzul11AdaptiveFastPaxosMakingQuickEverlasting}, Mencius
\cite{MaoJunqueiraMarzullo08MenciusBuildingEfficientReplicatedStateMachine},
Fast Mencius \cite{WeiETAL13FastMenciusMenciusLowCommitLatency},
the Chandra-Toueg algorithm
\cite{ChandraToueg96UnreliableFailureDetectorsReliableDistributedSystems}, the
Ben-Or algorithm \cite{BenOr83AnotherAdvantageFreeChoiceCompletelyAsynchronous},
and Raft \cite{OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

With the wide use of cloud computing, improvements in the performance and
efficiency of SMR algorithms have the potential to impact most online services:
advanced SMR algorithms can be implemented by cloud computing service providers
and benefit all their customer's applications.

However, devising new SMR algorithms, optimized with
particular situations or metrics in mind, is hard: today,
more than two decades after the first SMR algorithms
\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion} were proposed, devising new
SMR algorithms is still complex enough to justify a publication in a top conference
\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

The problem is that it is difficult to reuse existing algorithms and ideas because
SMR algorithms are not modular. When devising a new algorithm, all the careful
analysis, tests, and proofs have to be done again from scratch.

We propose a framework to make existing Generalized Consensus algorithms modular
and to enable adding new optimizations incrementally, without modification to
the existing. Generalized consensus, abbreviated GC, is a precise formulation of
State-Machine Replication that subsumes consensus. An algorithm in our framework
is composed of independent modules that can run, abort their execution, and pass
the baton to any other module on the fly.

Crucially, when adding a new module to an existing algorithm, all the
performance analysis, tests, and proofs of the existing modules remain valid and
only the new GC module needs to be analyzed: Our composition theorem guarantees
that the composition of any number of correct modules is a correct GC algorithm,
thus one can combine different GC algorithms, optimized for different situations
and metrics, without any changes and with guaranteed correctness. Our framework
therefore enables incremental development of GC algorithms, where each new
development can build on the previous ones instead of starting again from
scratch.

Moreover, our framework enables a clear separation of concerns which simplifies
the design of new individual modules. A composite GC algorithm built with our
framework works by running the GC module deemed the most appropriate at any
given time, dynamically changing GC module when needed. Therefore a composable
GC algorithm is not required to make progress in all situations: if it does
not make progress, the system can switch to another GC module. Thus algorithms
designers who want to create a new GC algorithm, optimized for particular
conditions, can focus only on the particular conditions of interest, knowing
that the other existing modules will handle the other cases.

The key novelty of our framework lies in its practical applicability:
\begin{itemize}
  \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
    that allows optimizing the execution of commuting commands.
  \item Many SMR algorithms are variations on a basic structure consisting of one of two kinds of 
    quorum systems \cite{GuerraouiVukolic10RefinedQuorumSystems}. 
    The first kind, that we call resilient quorums, are used in leader-based algorithms, while the second
    kind of quorums, that we call fast quorums, are used in leader-less algorithms and can order commuting commands without synchronization.
    We provide two specifications, at an intermediate level of abstraction, that capture the essence of composing algorithms using these two kinds of quorums.
    Our two specifications allow to easily check that new modules based on resilient or fast quorums are correct, covering a large spectrum of practical algorithms. 
\end{itemize}

\begin{comment}
Allowing the execution of commuting commands to proceed without
synchronization is an important practical optimization, as commands may
often commute in practice, for example when operating on disjoint parts
of a data structure. To enable the synchronization-free execution of
commuting commands, we have based our framework on Generalized Consensus
\cite{Lamport05GeneralizeConsensus}. Traditional SMR implementations use a
separate instance of a consensus algorithm to agree on the command to execute
at each index of the command sequence on which the replicas must agree. Using
separate consensus instances makes it hard to optimize the execution of
commuting command because commutativity is not a property of a single command.
Instead of restricting agreement to a single value, the Generalized Consensus
problem requires that set of replica servers agree on the state of a growing
data structure called a command-structure set. A command-structure set is
a data structure with an append operation such that the order in which two
commuting commands are appended does not change the resulting state. Thanks to
this property, two replicas executing commuting commands in a different order
will not end up in a different state and can continue executing without further
synchronization. In message-passing systems, this property can be leveraged to
reduce the latency of processing one command to two message delays instead of
three. 
%As we will see, fast execution of commutative requests also enables the processing of read-only commands
%in two message delays, although read-only commands do not commute with write
%commands.

In practice, a SMR algorithm must support fast reconfiguration to replace
faulty components, without which the long-term stability of a system is
inevitably compromised. Our framework abstracts over how the reconfiguration
decisions are made, providing an interface allowing to plug in an external
reconfiguration master implementing the reconfiguration policy of the
system. The master decides which GC module to run next and which replica
servers will participate, removing failed replicas and adding new ones. The
reconfiguration master enables tolerating $k$ failures with $k+1$ replicas
\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, an important
improvement over the traditional $2k+1$ needed replicas.
\end{comment}

To demonstrate the practical applicability of our framework we provide
composable versions of the generalized variations of popular consensus
algorithms: Classic Paxos, Fast Paxos, Chain Replication, and multi-coordinated
Paxos. Thanks to our specifications of algorithms based on quorum-systems, 
we have checked the correctness of our version of Fast Paxos and Chain Replication with the TLC
model-checker and we provide all the specifications needed to replicate these
experiments. Our framework is also fully formalized and its composition theorem
proved in Isabelle/HOL, a foundational interactive theorem prover, thereby
certifying the correctness of the composition theorem at the deepest level.
The TLA+ and Isabelle/HOL development can be found at \url{http://losa.fr/cgc}.

At the heart of our framework lies an abstraction, called Composable Generalized
Consensus (CGC), of the familiar notion of round (also called epoch, ballot, or
view in the literature) used in most SMR algorithms. As an example, let us
consider Fast Paxos, one of the first SMR algorithm which can change strategy
dynamically. Fast Paxos is a round-based SMR algorithm that has two types of
rounds: fast rounds and normal rounds. Fast rounds process requests faster
than normal rounds in contention-free executions but fall back to a slower
scheme under contention. Normal rounds are the same as in the classic Paxos
and are slower than fast rounds in contention-free executions. Any process can
initiate a round change at any point during execution: this allows switching
between fast and normal rounds as well as changing the leader from one normal
round to the next, similarly to classic Paxos. Each kind of round maintains
different invariants relating the state of each process and the values that
might have been chosen in the previous rounds, thus a given round must know
which past rounds are fast or normal rounds in order to take over the execution
safely. When adding more kinds of rounds such an approach will quickly become
unmanageable, as each type of rounds must have one case for each possible other
kind of round, making $O(n^2)$ cases to consider in total.

Our notion of Composable Generalized Consensus defines a common interface and a
common correctness property for a round, abstracting over the particular round
implementations and making all types of rounds compatible with each other.

Moreover, our interface for rounds enables a crucial optimization seen in
practical systems: introducing an external reconfiguration master which manages
the addition of new replicas and the removal of failed ones. Interestingly, as
explained in \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication},
in a real system it is often possible to implement a reliable reconfiguration
master which allows to tolerate $f$ faults with $f+1$ replicas instead of the
traditional $2f+1$ replicas. The reconfiguration master can also make the choice
of the new type of round to execute, providing the system's reconfiguration
policy.

>>>>

The State-Machine Replication technique (abbreviated SMR) allows building
reliable computing services on top of unreliable hardware. The basic idea is
to replicate a service over several servers and enforce, using a distributed
protocol, that each server executes the same sequence of deterministic commands.
In this fashion, a command will produce the same effect on all replicas.
Therefore, if a replica crashes, clients of the service can be transparently redirected to
another replica and will have the illusion of accessing a centralized service at
all times. SMR is often formally specified using the Generalized Consensus
specification~\cite{Lamport05GeneralizeConsensus}, abbreviated GC\@.

Because SMR is part of the basic infrastructure supporting cloud computing,
improving the performance of SMR algorithms can result in performance gains,
cost reduction, or improved resilience (because SMR may be used where it was
too costly before) for many online services.

However, devising new SMR algorithms or optimizing existing ones for a
particular use-case is hard: today, more than two decades after the first SMR
algorithms~\cite{Lamport98ParttimeParliament,BirmanJoseph87ReliableCommunicationPresenceFailures,OkiLiskov88ViewstampedReplicationGeneralPrimaryCopy,DworkLynchStockmeyer84ConsensusPresencePartialSynchronyPreliminaryVersion}
were proposed, devising a new SMR algorithm can justify a publication in a top
conference~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments,OngaroOusterhout14SearchUnderstandableConsensusAlgorithm}.

The problem is that it is difficult to reuse existing algorithms and ideas
because SMR algorithms are not modular. When devising a new algorithm, any
analysis, test, or correctness proof has to be done again from scratch.

To simplify the task of devising new SMR algorithms we propose a framework for
devising reusable SMR modules. A module in our framework is an SMR algorithm
augmented with an \emph{abort interface} for aborting its execution and passing
the baton to another module and with a matching \emph{init interface} for taking
over the execution of an aborting module. Two modules are composed by passing
the outputs produced by the abort interface of the first module, called abort
values, as inputs to the init interface of the second module, called init values
in the context of the second module.

A module must satisfy a correctness condition that we call Composable
Generalized Consensus (abbreviated CGC). Any correct module is a safe
Generalized Consensus~\cite{Lamport05GeneralizeConsensus} algorithm, but the progress guarantees may vary between
modules: CGC only requires a module to make progress executing commands or to
abort its execution. CGC also relates the init values, learned values (learned values represent the local state-machine state of individual replicas), and abort values in order for our composition theorem to hold. 
Generalized Consensus is a specification of SMR algorithms that takes into account the commutativity relation among commands and can off important performance advantages in practice.

The composition theorem enables incremental development of SMR algorithms
because it guarantees that any number of correct modules may be composed in
a sequence to form a correct SMR algorithm. Therefore, our framework allows
building \emph{dynamic SMR algorithms}, which dynamically change module when
the currently running module aborts, choosing the module best suited to the
current operating conditions. Thanks to the composition theorem, new modules can
be added to a dynamic SMR algorithm without modifying the existing modules and
with guaranteed correctness. Moreover, the relaxed progress requirements of CGC
enable a separation of concerns that simplifies optimizing SMR algorithms for
particular scenarios because there is no need to consider executing commands
under a situation outside the scenario: the module can just abort in this case.

In practice, we provide a TLA+ specification of CGC that the user of our framework can use to debug and prove new modules correct using the TLA$^+$ toolbox~\cite{TLATools}.
Moreover, we also provide two TLA+ specifications simplifying the development of modules based on Refined Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems}. These two specifications both implement CGC and provide a skeleton for a user to build her own module. The two specifications demonstrate how to update the local state of a replica after gathering the states or the init values of a write quorum of replicas, and how to
produce an abort value after gathering the state of a read quorum of replicas. 
The two specifications use different combinations of sizes of read and write quorums, which affects the invariants that the replicas must maintain and, in consequence, the latency with which a command can be processed but the module.
Our two specifications abstract over how to disseminate the commands to the replicas (e.g.\ whether to use a leader as in Paxos, multiple leaders as in MultiCoordinatedPaxos, or whether to broadcast commands without using a leader as in Fast Paxos).
The user can therefore specify a concrete command dissemination procedure and then model-check that the obtained algorithm refines one of the two specifications.  
We demonstrate the practical use of our framework by presenting modules based on Paxos~\cite{Lamport98ParttimeParliament}, Fast Paxos\cite{Lamport06FastPaxos}, Chain Replication~\cite{RenesseSchneider04ChainReplicationSupportingHighThroughputAvailability}, a case of Vertical Paxos~\cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication}, and EPaxos~\cite{MoraruAndersenKaminsky13ThereIsMoreConsensusEgalitarianParliaments}
(TODO, but would be nice).

Compared to related work, described in \cref{sec:related}, our key contributions are the following.
\begin{itemize}
    \item We provide tool support for designing CGC modules in the form of two TLA+~\cite{Lamport02SpecifyingSystems} specifications simplifying the design and model-checking of SMR algorithms based on Quorum Systems~\cite{GuerraouiVukolic10RefinedQuorumSystems} and demonstrate their use by designing CGC modules inspired by Paxos, Fast Paxos, and Primary-Backup replication.
    \item Our framework is based on Generalized Consensus, a generalization of the consensus problem 
        that allows optimizing the execution of commuting commands, a common practical case.
\end{itemize}
These two contributions set this paper apart from related work, notably the Abstract framework~\cite{GuerraouiETAL10Next700BftProtocols} and the Speculative Linearizability framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}, both described in \cref{sec:related}.
Our TLA+ specifications can be found at \url{http://losa.fr/cgc}.
Moreover, we have also formalized the CGC correctness condition and proved the composition theorem in Isabelle/HOL\@, and the Isabelle theories are available at the same URL\@.

### 2016-01-06 paper/paper.tex
Switched to the llncs template. Reworked the abstract and intro.

\input{header.tex}
\input{defs.tex}

@@

\begin{comment}
SMR is often implemented using a sequence of instances of Consensus.
There are numerous examples of consensus algorithms in the literature.
However, Composable Generalized Consensus is based on Generalized Consensus. 
In Generalized Consensus \cite{Lamport05GeneralizeConsensus}, the learners of a
SMR algorithm repeatedly learn new values belonging to a \emph{C-struct set}. A
C-struct can be seen as a data structure encoding  sequences of commands modulo
the ordering of commands that commute. The advantage of Generalized
Consensus over Consensus is that
two servers which receive commuting commands in different orders can still reach
agreement on C-structs without further synchronization. 
Our Composable Generalized Consensus can also use C-structs, with the same advantages
as Generalized Consensus. 
\end{comment}

Composable Generalized Consensus combines the ideas of the Abstract Framework \cite{GuerraouiETAL10Next700BftProtocols}, Generalized Consensus  \cite{Lamport05GeneralizeConsensus}, and Vertical Paxos \cite{LamportMalkhiZhou09VerticalPaxosPrimarybackupReplication} in a
unified abstraction.

@@

Our work builds on the Abstract Framework , which proposes to
build Byzantine Fault-Tolerant algorithms as the succession of abortable rounds
called Abstract instances. Abstract instances must satisfy the Abstract
correctness properties. By construction, the composition of any number of
Abstract instances is a correct SMR implementation. The Abstract
Framework is similar to our framework but uses totally ordered sequence of commands, making it difficult apply it to 
practical algorithms that optimize the execution of commuting and read-only commands.

Finally, the Speculative Linearizability framework
\cite{GuerraouiKuncakLosa12SpeculativeLinearizability} was a first
attempt by the authors of the present paper at generalizing the
Abstract Framework and mechanically proving its properties with the
Isabelle/HOL interactive proofs assistant.  Like the Abstract Framework, the
Speculative Linearizability framework is based on totally ordered
sequences of commands, with the same drawback for the execution of
commuting and read-only commands. Moreover, Composable Generalized Consensus is both
conceptually simpler and subsumes the Speculative Linearizability
property. Composable Generalized Consensus
and its main property, the composition theorem, are also formalized in Isabelle/HOL.

\begin{comment}
Composable Generalized Consensus is the most complete framework to date for building
adaptive replication algorithms out of heterogeneous rounds, making state of the
art optimization techniques available to algorithm designers, which was not
possible before.
\end{comment}

@@

We consider a set of asynchronous servers communicating by message passing in
an asynchronous network. Servers are not necessarily sequential processes.
We describe a behavior of the system using a sequence of states,
where a state is a function from variable names to values (i.e. a valuation of
variables). We specify sets of allowed behaviors using an initial predicate
and a next-state relation expressed using unprimed variables and primed variables (next-state variables) in the
style of TLA+ \cite{Lamport02SpecifyingSystems}. However, for succinctness, our
specification are rather informal, eluding non-essential details.

@@

The Broadcast-Decide algorithm can process a proposal in two message delays when there are no conflicts, i.e. no contention on non-commuting commands. However, if there are repeated conflicts, the algorithm may not make progress despite changing round, as the acceptors of the each new round
may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to learn new histories under high contention by
designating a leader server that will totally order the proposals. However, to make our Broadcast-Decide algorithm work like Fast Paxos, we would have to modify it substantially. If we had spent effort testing Broadcast-Decide implementations and proving it correct, this effort would be wasted and we would have to redo all the tests and proofs from scratch.

@@

\begin{thm}[Composition Theorem]
    \label{thm:comp}
Consider the composition of two rounds as defined above.
Define the state variable $proposed$ as $proposed_1 \cup proposed_2$. For every learner $l$, define 
$learned\left[ l \right]$ as the maximum of $learned_1\left[ l \right]$ and $learned_2\left[ l \right]$.
For every switcher $s$, define $inits\left[ s \right]$ as $inits_1\left[ s \right]$ and $aborts\left[ s \right]$ as $aborts_2\left[ s \right]$.

Then the sequence of states of the variables $proposed$, $learned$, $inits$, and $aborts$ satisfy the specification of Composable Generalized Consensus.
\end{thm}

@@

We provide a mechanically-checked proof of the composition theorem in Isabelle/HOL at \url{http://losa.fr/cgc}.
The proofs uses a version of the specification of CGC that has a more
executable form that presented in this section. 
The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
is also formalized in Isabelle/HOL. The proof uses a refinement
mapping from the composition of two consecutive CGC instances to a single CGC
instance.

\begin{comment}
\begin{thm}[Composition Theorem]
    \label{thm:comp}
    In every execution of a succession of CGC instances, the history of valuations
    of the variable $proposed$ and, for every learner $l$, of the variables
    $learned\left[ l \right]$ satisfies Generalized Consensus.
\end{thm}
\Cref{thm:comp} can be shown by induction on the number of rounds,
using the following lemma.
\begin{lem}
  \label{lem:comp}
    Consider two rounds numbered $i$ and $i+1$. In every execution of a
    succession of CGC instances, the history of valuations of the variables $inits
    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies CGC.
\end{lem}
\begin{proof}
  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
  The proofs uses a version of the specification of CGC that has a more
  executable form that presented in this section. Moreover,
  instead of history of commands, the Isabelle/HOL specification uses the more
  general notion of command history, presented in the next subsection.
  The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
  is also formalized in the Isabelle/HOL. The proof uses a refinement
  mapping from the composition of two consecutive CGC instances to a single CGC
  instance.
\end{proof}

@@

Fast and resilient CGC round are two specifications, at an
intermediate level of abstraction, which satisfy CGC. Both
specifications describe how the state of the acceptors should be
updated, using the notion of quorum, but abstract over the particular communication topology used
to implement the state accesses: the specifications use a global state
which is accessible by all servers in all roles. One can refine fast
or resilient rounds by implementing the state accesses and coordination
using the network, obtaining a concrete algorithm.

@@

Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
updated each time a new local history is accepted by the learner $l$ by
inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
definition of a history variable).
\begin{thm}
  \label{thm:rounds}
    Both resilient and fast CGC round implement CGC under the refinement
    mapping that consists in substituting $\overline{learned\left[ l \right]}$
    for $learned\left[ l \right]$.
\end{thm}

@@

\begin{thm}

@@

\end{thm}

>>>>

\documentclass{llncs}
\usepackage[english]{babel}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{comment}
\usepackage{datetime}
\usepackage[url=true,doi=false,isbn=false,backend=biber]{biblatex}
\usepackage{mathtools}
\usepackage{paralist}
\usepackage{mathtools}

%\input{header.tex}
\input{defs.tex}

@@

Our work is closely related to the Abstract
framework~\cite{GuerraouiETAL10Next700BftProtocols}
and the Speculative Linearizability
framework~\cite{GuerraouiKuncakLosa12SpeculativeLinearizability}. The
Abstract framework pioneered the idea of decomposing SMR algorithms in
independent modules and experimentally showed that the resulting composite
algorithms outperformed state-of-the-art monolithic algorithms. Compared
to our work, the Abstract framework is not precisely formalized and lacks
tools supporting the development of correct algorithms in the framework,
which are direly needed given the notorious complexity of SMR algorithms.
Moreover, the Abstract framework does not allow optimizing for the execution
of commuting commands. However, the Abstract framework supports byzantine
fault-tolerant~\cite{LamportShostakPease82ByzantineGeneralsProblem} SMR
algorithms, which we do not.

The Speculative Linearizability framework was a
first attempts by the authors to formalize the Abstract framework, provide
formal proof support in Isabelle/HOL~\cite{NipkowPaulsonWenzel02IsabelleHOL},
and extend the Abstract framework to support algorithms that optimize
the execution of commuting commands by requiring Speculatively
Linearizable modules to refine Abstract modules (under a forward/backward
simulation~\cite{LynchVaandrager95ForwardBackwardSimulationsIUntimedSystems})
modulo an equivalence relation on histories of commands that reflects the
commutativity relation among commands. However, the simulation technique led to
complex correctness statements that were hard to scale beyond the examples of
the paper, and no realistic protocol was built in the framework. The present
work addresses the shortcoming of the Speculatively Linearizable framework by
using the notion of command-structure set introduced for the definition of
Generalized Consensus~\cite{Lamport05GeneralizeConsensus}. Moreover, we show how
several well known and practical algorithms can be cast and optimized in our
framework, providing a set of TLA+ specifications that ease the design and model
checking of algorithms based on quorum systems.

TODO\@: expand the following.

@@

@@

We consider a set of asynchronous servers communicating by message passing in
an asynchronous network. Servers are not necessarily sequential processes.
We describe a behavior of the system using a sequence of states,
where a state is a function from variable names to values (i.e.\ a valuation of
variables). We specify sets of allowed behaviors using an initial predicate
and a next-state relation expressed using unprimed variables and primed variables (next-state variables) in the
style of TLA+ \cite{Lamport02SpecifyingSystems}. However, for succinctness, our
specification are rather informal, eluding non-essential details.

@@

The Broadcast-Decide algorithm can process a proposal in two message delays when there are no conflicts, i.e.\ no contention on non-commuting commands. However, if there are repeated conflicts, the algorithm may not make progress despite changing round, as the acceptors of the each new round
may be initialized with incompatible histories. The Fast Paxos algorithm \cite{Lamport06FastPaxos} is able to learn new histories under high contention by
designating a leader server that will totally order the proposals. However, to make our Broadcast-Decide algorithm work like Fast Paxos, we would have to modify it substantially. If we had spent effort testing Broadcast-Decide implementations and proving it correct, this effort would be wasted and we would have to redo all the tests and proofs from scratch.

@@

\begin{theorem}[Composition Theorem]
    \label{thm:comp}
Consider the composition of two rounds as defined above.
Define the state variable $proposed$ as $proposed_1 \cup proposed_2$. For every learner $l$, define 
$learned\left[ l \right]$ as the maximum of $learned_1\left[ l \right]$ and $learned_2\left[ l \right]$.
For every switcher $s$, define $inits\left[ s \right]$ as $inits_1\left[ s \right]$ and $aborts\left[ s \right]$ as $aborts_2\left[ s \right]$.

Then the sequence of states of the variables $proposed$, $learned$, $inits$, and $aborts$ satisfy the specification of Composable Generalized Consensus.
\end{theorem}

@@

We provide a mechanically-checked proof of the composition theorem in Isabelle/HOL at \url{http://losa.fr/cgc}.
The proofs uses a version of the specification of CGC that has a more
executable form that presented in this section. 
The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
is also formalized in Isabelle/HOL\@. The proof uses a refinement
mapping from the composition of two consecutive CGC instances to a single CGC
instance.

\begin{comment}
\begin{theorem}[Composition Theorem]
    \label{thm:comp}
    In every execution of a succession of CGC instances, the history of valuations
    of the variable $proposed$ and, for every learner $l$, of the variables
    $learned\left[ l \right]$ satisfies Generalized Consensus.
\end{theorem}
\Cref{thm:comp} can be shown by induction on the number of rounds,
using the following lemma.
\begin{lem}
  \label{lem:comp}
    Consider two rounds numbered $i$ and $i+1$. In every execution of a
    succession of CGC instances, the history of valuations of the variables $inits
    = inits_i$, $propose = proposed_i \cup propose_{i+1}$, for every learner
    $l$, $learned\left[ l \right] = learned_i\left[ l \right] \cup
    learned_{i+1}\left[ l \right]$and $aborts = aborts_{i+1}$ satisfies CGC.
\end{lem}
\begin{proof}
  A mechanically-checked proof in Isabelle/HOL appears in \cref{sec:isaproofs}.
  The proofs uses a version of the specification of CGC that has a more
  executable form that presented in this section. Moreover,
  instead of history of commands, the Isabelle/HOL specification uses the more
  general notion of command history, presented in the next subsection.
  The Isabelle/HOL formalization of CGC uses the theory of I/O automata, which
  is also formalized in the Isabelle/HOL\@. The proof uses a refinement
  mapping from the composition of two consecutive CGC instances to a single CGC
  instance.
\end{proof}

@@

Fast and resilient CGC round are two specifications, at an
intermediate level of abstraction, which satisfy CGC\@. Both
specifications describe how the state of the acceptors should be
updated, using the notion of quorum, but abstract over the particular communication topology used
to implement the state accesses: the specifications use a global state
which is accessible by all servers in all roles. One can refine fast
or resilient rounds by implementing the state accesses and coordination
using the network, obtaining a concrete algorithm.

@@

Define, for every learner $l$, the history variable $\overline{learned}\left[ l \right]$ which is initialized to the empty set and which is
updated each time a new local history is accepted by the learner $l$ by
inserting the new local history in $\overline{learned}\left[ l \right]$ (see \cite{AbadiLamport91ExistenceRefinementMappings} for the
definition of a history variable).
\begin{theorem}
  \label{thm:rounds}
    Both resilient and fast CGC round implement CGC under the refinement
    mapping that consists in substituting $\overline{learned\left[ l \right]}$
    for $learned\left[ l \right]$.
\end{theorem}

@@

\begin{theorem}

@@

\end{theorem}
